\subsection{Conclusions}
\label{sec:sec_conc}
In this article, we have presented a novel distributed processing framework, which supports a probabilistic classification and a regression approach. Its core algorithm is an extension of a distributed approximate $k$NN implementation, optimized to operate efficiently in a single distributed session. It was implemented with the Apache Flink scalable data processing engine.

We have conducted a detailed experimental evaluation in order to assess the framework in terms of wall-clock completion time and scalability, comparing it with similar approaches based on Apache Spark and Apache Hadoop. Our framework outperformed all competing implementations, managing to execute significantly faster on the same workloads and scale better for larger datasets, as a result of our optimizations and its ability to execute in a single distributed session. Furthermore, we performed two water consumption related real-world case studies, demonstrating our framework's potential in knowledge extraction tasks on data of very high volume.

With real-world data collection continuously evolving in all fields, knowledge extraction from daily activities becomes very challenging. The directions for future work are as follows. We will perform extended case studies on more datasets from various sources, in order to establish our framework's ability in performing ad-hoc data mining tasks. Furthermore, we will explore its applicability on data stream mining applications, where the input is a continuous flow of data records. Finally, we will enhance our framework's knowledge discovery capacity, by extending it with more distributed machine learning approaches, in an attempt to raise its potential on the continuously growing field of Big Data analytics.